Competing Bandits: \\ The Perils of Exploration under Competition

### 2018-11-05 paper/content/introduction.tex
superficial tweaks to intro

We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}. \swcomment{confused about how they correspond to ``competition intensity''}Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}. \gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. \gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that perform better in the long run. As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}

@@

The study of whether data can give a significant \swdelete{competitive} advantage
has been discussed theoretically in \cite{varian2018artificial,
  lambrecht2015can} and empirically in
\cite{bajari2018impact}. \swedit{While they argue that small amounts
  of additional data do not provide significant improvement in
  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.
\swdelete{
These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers. In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm. Additionally, the first-mover advantage has been well-studied in economics and marketing \cite{kerin1992first}. }

>>>>

We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}. \swcomment{confused about how they correspond to ``competition intensity''}Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}. \gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. \gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that performs better in the long run. As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}

@@

The study of whether data can give a significant \swdelete{competitive} advantage
has been discussed theoretically in \cite{varian2018artificial,
  lambrecht2015can} and empirically in
\cite{bajari2018impact}. \swedit{While they argue that small amounts
  of additional data do not provide significant improvement in
  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.
\swdelete{
These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers. In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm. Additionally, the first-mover advantage has been well-studied in other settings in economics and marketing \cite{kerin1992first}. }
